---
title: "HW 3: Distributional consequences of climate policy"
author: "Grace Bianchi, Claire Meuter, and Katheryn Moya"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, options(scipen=999))
library(here)
library(tidyverse)
library(equatiomatic)
library(broom)
library(janitor)
library(rootSolve)
library(PlaneGeometry)
library(tibble)
```
Provided Information:

- Current electricity cost is $0.10/kWh
- One kWh electricity emits 0.85 lbs CO^2^
- 2204.62 lbs in 1 metric ton 
- Biden's Interim SCC value of $51/metric ton or 5100 cents/metric ton (total external cost)

```{r}
# read in data
energy_demand <- read_csv("HW3_data.csv") %>% 
  select(-1) %>% 
  clean_names()
```

## Question 1
One kWh of electricity emits 0.85 pounds of CO^2^. Assuming that the interim SCC correctly reflects the total social cost of one metric ton of CO2, what is the marginal externality cost per kwH of electricity?

```{r}
# convert lbs to metric ton
lb_to_ton <- (0.85/2204.62) 
# one kWh electricity emits 0.0003856 metric tons of carbon
MEC <- 5100*lb_to_ton # $.0197
``` 

- **MEC = $ `r round((MEC/100), 4)` cents/kWh electricity**

## Question 2

What is the aggregate monthly demand curve for electricity? What is the supply curve for electricity? What is the “benefit” to consumers under the status quo? What is the “benefit” to producers under the status quo? What is the environmental cost under the status quo?

```{r}
# define function to get demand curve
low_income <- lm(price_cents ~ q_low_kwh, data = energy_demand)
high_income <- lm(price_cents ~ q_high_kwh, data = energy_demand)

demand <-function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)  #addresses differences in y-intercepts
  return(q)
}

# define aggregate demand function
demand_agg <- function(p){
  q <- demand(p, low_income) + demand(p, high_income)
  return(q)
}

price = seq(0, 30, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()
agg_energy_demand<- tibble(Qagg = Qagg, price = price)
market_q <- demand_agg(10)

supply_slope <- (10/market_q)
supply <- function(q){
  p <- supply_slope * q
  return (p)
}
```

```{r}
# create demand and supply curve
ggplot() +
  geom_line(agg_energy_demand, mapping = aes(x = Qagg, y =price, color = "agg. Demand")) +# Agg demand
  #geom_line(agg_energy_demand_tax, mapping = aes(x=Qagg_tax, y = price_tax, color = "taxed agg. demand")) + #taxed agg demand
  geom_abline(slope = coef(low_income)[["q_low_kwh"]], 
              intercept = coef(low_income)[["(Intercept)"]], color="purple", linetype="twodash") + # Low consumer demand
  geom_abline(slope = coef(high_income)[["q_high_kwh"]], 
              intercept = coef(high_income)[["(Intercept)"]], color="steelblue", linetype="twodash") + # High consumer demand
  geom_abline(slope = (10/market_q), intercept = 0, color = "orange", aes(value = "supply")) + # supply curve
  geom_abline(slope = 0, intercept = MEC, color = "darkolivegreen3", aes(value = "MEC")) + # MEC
  scale_x_continuous(limits = c(0,850000), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0,35), expand = c(0, 0)) +
  labs(x = "Electricity (kWh)", y = "Price (cents)") +
  #geom_line(taxcurve, mapping = aes(x = Qagg, y = price, color = "orange"))+
  theme_bw() +
  scale_color_manual(breaks = c("agg. demand", "supply", "MEC", "low consumer demand", "high consumer demand"),
                     values = c("agg. demand" = "black", "supply" = "orange", "MEC" = "darkolivegreen3", "high consumer demand" = "steelblue", "low consumer demand" = "purple"))

```

The aggregate monthly demand curve (black line) is the horizontal sum of the low income and the high income demand curves (dashed lines). The supply or MPC curve (orange line) is P = `r supply_slope` Q


```{r}
# consumer surplus for low and high income
CS <- function(p, model) {
  q <- demand(p, model)
  cs <- 0.5 *(model$coefficient[[1]]- p)*q
  return(cs)
}
low_income_CS <- CS(10, low_income) / 100
high_income_CS <- CS(10, high_income) /100

# consumer surplus of agg. demand
CS_agg <- function(p){
  cs <- CS(p, low_income) + CS(p, high_income) # should low_income_cs be used here instead of initial demand
  return(cs)
}
# plug in P* and divide by 100 to get $
CS_total <- CS_agg(p=10) / 100

# producer surplus
PS <- function(q){
  0.5*supply(q) *q
}

PS_total <- PS(market_q) / 100

env_cost <- (MEC * market_q)/100

```


Under the status quo:

- **Consumer benefit = $`r round(CS_total, 2)` **
- **Producer benefit = $`r round(PS_total, 2)` **
- **Environmental cost = $`r round(env_cost, 2)` **

## Question 3
How is the current consumer benefit divided between “high” and “low” income consumers?

**The high-income consumers receive $ `r round(high_income_CS, 2)` and the
low-income consumers receive $`r round(low_income_CS, 2)`**

## Question 4 
Derive the optimal electricity tax (in cents per kWh) using the interim SCC. Noting that recent research has shown the poor face a disproportionate share of the impacts from climate change, assume that the climate externality is borne entirely by the “low” income group. 
- 2204.62 lbs in 1 metric ton 
- Interim SCC value of $51/metric ton or 5100 cents/metric ton (total external cost)

```{r}
# new aggregate with tax 
demand_tax <- function(p, model){
  q <- (p - (model$coefficients[[1]]- MEC ))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

demand_agg_tax <- function(p){
  q <- demand_tax(p, low_income) + demand_tax(p, high_income) #should this be demand_tax(p,high_income)?
  return(q)
}

demand_high_taxed <- function(p,t){
  q <-demand_tax(p, high_income,MEC)
  return(q)
}
demand_low_taxed <- function(p,t){
  q <- demand_tax(p, high_income,MEC)
  return(q)
}

price_tax = seq(0, 30, length.out = 100)
Qagg_tax <- map(price_tax, demand_agg_tax) %>% unlist()

agg_energy_demand_tax<- tibble(Qagg = Qagg_tax, price = price_tax)

market_p_tax <- demand_agg_tax(10) # P*
#plug in market_q_tax into new demand tax equation
market_q_tax <- demand_agg_tax(10)

```


```{r}
supply_pointx <- 15/supply_slope

tax_demand_line <- Line$new(B = c(801868.98,0), A = c(235880.43,20))
supply_line <- Line$new(A = c(0,0), B = c(supply_pointx, 15))
tax_intersect <- intersectionLineLine(tax_demand_line, supply_line)
# 525035.350330      9.782305
```

(a) The amount of electricity produced and consumed is **`r round(tax_intersect[1], 2)` kWh ** 
(b) The electricity price is **`r round(tax_intersect[2], 2)` cents per kWh**

```{r}
tax_pstar <-tax_intersect[2] #9.78
tax_qstar <- tax_intersect[1] 

# Consumer surplus calculations
high_sum <- summary.lm(high_income)
high_yint <- (high_sum$coefficients[1]) - MEC #33.57
CS_high_tax <- 0.5*tax_qstar*((high_yint-tax_pstar)) 

low_sum <- summary.lm(low_income)
low_yint <- (low_sum$coefficients[1]) - MEC

CS_low_tax <- 0.5*tax_qstar*((low_yint) - (tax_pstar - MEC))
CS_low_tax_ec <- CS_low_tax - (env_cost)
# test
```
(c) Overall welfare of "high" income consumers: **$ `r round((CS_high_tax/100), 2)` ** NOT CORRECT
(d) Overall welfare of "low" income consumers: **$ `r round((CS_low_tax_ec/100), 2)` ** NEED TO FIX

```{r}
power_suppliers <- PS(tax_qstar) /100
```

(e) Power suppliers: **$`r round(power_suppliers, 2)`**

```{r}
total_env_d <- (MEC*tax_qstar)/100
```

(f) Total environmental damage: **$`r round(total_env_d, 2)`**

```{r}
tax_revenue <- MEC * tax_qstar
```

(g) Total tax revenue generated: **$`r round((tax_revenue/100), 2)`**

## Question 5
Now, assume that all revenue from the electricity tax will be redistributed to the consumers in proportion to their pre-tax consumption. For example, if 80% of the electricity was consumed by “high” income consumers, then they get 80% of the tax revenue. Additionally, consider the fact that current scientific evidence suggests the true SCC may be much higher than $51. For a range of SCC values ($51, $75, $100, $125, and $150 per metric ton of CO2), calculate the effects of an SCC-based electricity tax on:

```{r, eval = FALSE}
# percent tax revenue for each income group
high_percent_consumption <- demand(tax_pstar, high_income)/tax_qstar
low_percent_consumption <- 1 - high_percent_consumption
# tax revenue for each group in $
high_share <- tax_revenue * high_percent_consumption
low_share <- tax_revenue * low_percent_consumption

MEC_51 <- 5100 * lb_to_ton
MEC_75 <- 7500 * lb_to_ton
MEC_100 <- 100000 * lb_to_ton
MEC_125 <- 125000 * lb_to_ton
MEC_150 <- 150000 * lb_to_ton

#create new dataframe for tax effects table
SCC <- c(51,75,100,125,150)

# new supply price function 
supply_price <- function(q){
  p <- supply_slope *q 
  return(p)
}

#manipulate data for table
cost_table <- tibble(SCC) %>% 
  mutate(MEC = SCC * lb_to_ton) %>% 
  mutate(q_star = demand_agg_tax(MEC)) %>% 
  mutate(p_star = supply_price(tax_qstar)) %>% 
  mutate(p_star_tax = p_star + MEC) %>% 
  mutate(tax_revenue = q_star * MEC) %>% 
  mutate(high_cs_rev = high_percent_consumption * tax_revenue) %>% 
  mutate(low_cs_rev = low_percent_consumption * tax_revenue) %>% 
  mutate(cs_high_without_rev = CS(p_star_tax, high_income)) %>% 
  mutate(cs_high_with_rev = (cs_high_with_rev +high_cs_rev)/100) %>%
  mutate(cs_low_without_rev = CS(p_star_tax, low_income) - (MEC * tax_qstar)) %>%
  mutate(cs_low_with_rev = (cs_low_without_rev + low_cs_rev)/100) %>% 
  mutate(ps = PS(tax_pstar)/100) %>% 
  mutate(TEC = (MEC * tax_qstar))

#create table by filtering data of interest
filtered_cost = subset(cost_table, 
                       select = c(SCC, cs_low_with_rev, cs_high_with_rev,ps))

kable(filtered_cost, caption = "Table 1. Distribution Impacts of Social Costs of carbon among high and low income consumers and producers", 
      col.names = c("Social Cost of Carbon ($/ metric ton)", "High Income Welfare ($)", "Low Income Welfare ($)", "Electricity Producer Welfare ($)")) 


```

(a) Overall welfare of “high” income consumers

```{r}
high_CS_mec51 <- 0.5 * CS_high_tax
```

(b) Overall welfare of “low” income consumers

(c) Electricity producers

## Question 6

6.Suppose the “high” income group has access to expensive home solar generation. This lowers the electricity demand curve for the “high” income group by half (vertically) (aka by price). Under this new demand:


```{r}
#new demand curve lowered
demand_half <-function(p, model){
  q <- ((p*0.5) - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)  #addresses differences in y-intercepts
  return(q)
}

# define aggregate demand function for new half demand- note that high income gets new half demand where low income keeps same demand
demand_agg_num6 <- function(p){
  q <- demand(p, low_income) + demand_half(p, high_income)
  return(q)
}

#making a price vector
price = seq(0, 30, length.out = 100)
#applied the agg demand function to the price vector
Qagg_num6 <- map(price, demand_agg) %>% unlist()
#new dataframe with Qaggnum6 and price
agg_energy_demand_num6<- tibble(Qagg = Qagg_num6, price = price)


intercept_solar <- uniroot(function(p)
  demand_agg_num6(p) - supply(p),
  interval = c(0,100))


p_solar <- intercept_solar[["root"]]

q_solar <- 242802.5

Qagg_solar <- map(price, demand_agg_num6) %>% unlist()

agg_demand_solar <-tibble(Qagg = Qagg_solar, price = price)

```

(a) The total electricity consumption is **`r round(q_solar, 4)` kWh ** 
 

```{r}
TEC_solar <- (q_solar* MEC)/100 # solar_Q * MEC
```

(b)The total environmental externality is $**`r round(TEC_solar, 2)`** 


```{r}
environmental_tax <- (MEC/100)/(TEC_solar/total_env_d)
```

(c) What value of the electricity tax makes the total environmental damage the same as the damage when solar panels are available to the high income group? The value of the electricity tax that makes the total environmental damage the same as the damage when solar panels are available to the high income group is **`r (environmental_tax)` kWh **