---
title: "ESM204_HW3"
author: "Katheryn Moya"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(broom)
library(janitor)
library(rootSolve)
```

to create a function, we want to sum Qs
step one is to define function to set a demand
function is basically saying that first co

```{r}
# read in data
energy_demand <- read_csv("HW3_data.csv") %>% 
  select(-1) %>% 
  clean_names()
```

```{r}
# linear regression model

low_income <- lm(price_cents ~ q_low_kwh, data = energy_demand)
high_income <- lm(price_cents ~ q_high_kwh, data = energy_demand)

```

```{r}
# define function to get demand curve

demand <-function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)  #addresses differences in y-intercepts
  return(q)
}

#aggregate demand curve
demand_agg <- function (p) {
  q <- demand(p, low_income) + demand(p, high_income) 
}

price = seq(0,30,length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()
```


```{r}
# draw curves
agg_demand <- tibble(Qagg = Qagg, price = price)
low_demand <- tibble(q_low_kwh = q_low_kwh , price = price)
high_demand <- tibble(q_high_kwh = q_high_kwh , price = price)

ggplot(agg_demand, aes(Qagg, price)) +
  geom_line() 

+
  geom_line(aes(y = q_low_kwh), color = "green") +
  geom_line(aes(y = q_high_kwh), color = "blue")

  
  
  
  
  
  geom_abline(slope = coef(low_income)[["q_low_kwh"]], 
              intercept = coef(low_income)[["(Intercept)"]], color="green") + # low income demand
  geom_abline(slope = coef(high_income)[["q_high_kwh"]], 
              intercept = coef(high_income)[["(Intercept)"]], color="purple") + # high income demand
  scale_x_continuous(limits = c(0,850000), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0,35), expand = c(0, 0)) +
  labs(x = "Electricity (kWh)", y = "Price (cents)") +
  theme_bw()
```




## Question 1
- Current electricity cost is $0.10/kWh
- One kWh electricity emits 0.85 lbs CO2
- 2204.62 lbs in 1 metric ton 
- Interim SCC value of $51/metric ton or 5100 cents/metric ton (total external cost)

```{r}
# convert lbs to metric ton
lb_to_ton <- (0.85/2204.62) 
# one kWh electricity emits 0.0003856 metric tons of carbon
MEC <- 5100*lb_to_ton # $.0197
``` 

- MEC = $ `r round((MEC/100), 2)` cents per kWh electricity

## Question 2

```{r}
model_demand_l <- lm(price_cents  ~ q_low_kwh, data=energy_demand)
model_demand_h <- lm(price_cents ~ q_high_kwh, data=energy_demand)
# need to rearrange the parameter to get Q(P)! 
# Qagg = Qlow(P) + Qlow(h) 
# Importantly, since their intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take care of the kink.
# define a function to get demand
demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
# for each p level, return estimated aggregate demand
demand_agg <- function(p){
  q <- demand(p, model_demand_l) + demand(p, model_demand_h)
  return(q)
}
price = seq(0, 31.6, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()
df2<- tibble(Qagg = Qagg, price = price)
market_q <- demand_agg(10)
#taxcurve <- df2 - MEC
# new aggregate with tax 
demand_tax <- function(p, model){
  q <- (p - (model$coefficients[[1]]-1.97))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
demand_agg_tax <- function(p){
  q <- demand_tax(p, model_demand_l) + demand(p, model_demand_h) # high consumers are not subjugated to the tax
  return(q)
}
price_tax = seq(0, 31.6, length.out = 100)
Qagg_tax <- map(price, demand_agg_tax) %>% unlist()
df2_tax<- tibble(Qagg = Qagg_tax, price = price_tax)
market_q_tax <- demand_agg_tax(10)
ggplot() +
  geom_line(df2, mapping = aes(x = Qagg, y =price, color = "black")) +# Agg demand
  geom_line(df2_tax, mapping = aes(x=Qagg_tax, y = price_tax)) + #taxed agg demand
    geom_abline(slope = coef(model_demand_l)[["q_low_kwh"]], 
              intercept = coef(model_demand_l)[["(Intercept)"]], color="goldenrod", linetype="twodash") + # Low consumer demand
  geom_abline(slope = coef(model_demand_h)[["q_high_kwh"]], 
              intercept = coef(model_demand_h)[["(Intercept)"]], color="steelblue", linetype="twodash") + # High consumer demand
  geom_abline(slope = (10/market_q), intercept = 0, color = "green") + # supply curve
  geom_abline(slope = 0, intercept = MEC, color = "red") + # MEC
  #geom_abline(slope = (10/market_q), intercept = MEC, color = "purple") + #we need to delete this line because this is a tax on the consumer
  scale_x_continuous(limits = c(0,850000), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0,35), expand = c(0, 0)) +
  labs(x = "Electricity (kWh)", y = "Price (cents)") +
  #geom_line(taxcurve, mapping = aes(x = Qagg, y = price, color = "orange"))+
  theme_bw()
slope_supply <- (10/market_q)
# supply curve equation: P = 0.0000186Q
```

```{r}
# calculating surpluses at market equilibrium
# CS calculation needs Y intercept of high demand/demand agg curve --> at 0 kWh, price would be $0.316
high_sum <-summary.lm(model_demand_h)
high_y_int <- high_sum$coefficients[1, 1] 
low_sum <- summary.lm(model_demand_l)
low_y_int <- low_sum$coefficients[1, 1]
# Converted back to dollars
cs_sq_high <- 0.5*market_q*((high_y_int/100) - 0.10)
# High consumer benefit $57,984.10
cs_sq_low <- 0.5*market_q*((low_y_int/100) - 0.10)
# Low consumer benefit $35,882.29
cs_sq_total <- cs_sq_high + cs_sq_low
# Consumer benefit $93,866.39
ps_sq <- 0.5*market_q*0.10
# Producer benefit $26,835.97
env_cost_sq <- (MEC*market_q)/100
# Environmental cost $10,553.65
```

- Aggregate monthly demand curve is the black curve with a kink, which is the horizontal sum of the low (gold dashed) and high (blue dashed) demand curves.
- The supply curve is the green curve, which represents the marginal private cost curve. Its equation is P = `r round(slope_supply, 7)`Q

Under status quo:
- Consumer benefit = $`r round(cs_sq_total, 2)` 
- Producer benefit = $`r round(ps_sq, 2)`
- Environmental cost = $`r round(env_cost_sq, 2)`

## Question 3

The high-income consumers get much more of total benefit, at `r round(cs_sq_high, 2)` USD compared to `r round(cs_sq_low, 2)`USD for low-income consumers.

## Question 4 

